Realistically Interesting Domains for Probabilistic Planners

### 2014-04-07 FinalPaper/paper.tex
Add section heading for the domain evaluation
rewrote everything excepr gourmand

@@

\section{FF-Replan}
\label{sec:planners}

At the first international probabilistic planning competition IPPC-04, most entries were probabilistic planner using Markov Decision Processes. The winner, FF-Replan \cite{FFReplan}, however was based on deterministic planning techniques. FF-Replan is an action selection algorithm for online planning in probabilistic domains which deals with uncertainty by replanning. FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of the planning. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state mapped in an initially empty hash-table. This plan is executed until an unexpected state, not available in the hash-table is encountered.

@@

FF-Replan has some obvious shortcomings.Iit first determinizes the input domain and thus removes all probabilistic information from the problem. The success of the planner is therefore fully reliable on the chose determinization heuristic. As FF-replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-replan performs poorly in domains containing dead-ends. However, by relying on search FF-replan functions with enormous speed and efficiency and caused the International Conference on Automated Planning and Scheduling (ICAPS) to change their benchmarks to increase their \emph{probabilistically interestingness}. By evaluating and correcting the IPPC-11 domains on their \emph{realistically interestingness} we will draw conclusions whether or not these adjusted realistic domains are actually classical planning problems in disguise - ideal planning problems for FF-replan.

@@

\subsection{GOURMAND}

The IPPC-11 introduces more realistic MDPs to accommodate real-world systems with large branching factors cause by \emph{natural dynamics} and a limited time for the planner to come up with a policy. For this reason planners needed to be scalable and sensitive to the expected reward maximization. Glutton introduces an off-line algorithm based on LR\textsuperscript{2}TDP and relies on the fact that a successful policy can be produced by solving an MDP $M(H)$ for a smaller horizon $h$. Incrementally solving the sequence $M(1)$, $M(2)$ forms the basis of LR\textsuperscript{2}TDP\@. By iterating on the distance from the horizon, data can be reused (and is stored in the cache until the hash-table is full). The max-reward heuristic finds the largest horizon value $h'$ for which already a heuristic value $V$ is estimated. The  computation time remains much lower because it runs quickly into states that converged while solving $M(h-1)$. The max-reward computation relies on knowing the maximum $R_{max}$ any action can yield in any state or an upper bound on it.

\subsubsection{LRTDP}
Using goal oriented MDPs, the \emph{Labelled Real-Time Dynamic Programmer} (LRTDP) operates a series of trials starting from $s_0$.

\begin{itemize}
	\item Trials start at $s_0$
	\item Choose best action greedily in the current state based on current $V$
	\item Perform  Bellman backup on the current state
	\item Sample the outcome as the chosen action
	\item Transition to the new state
	\item Repeat until a goal, dead end or converged state is reached
	\item After trial, special convergence check on all states in the trial is performed to prove the convergence of these values
\end{itemize}

\subsubsection{LRTDP-FH}

Adapting LRTDP to a finite horizon MDP \emph{M(H)} requires to let each trial start at ($s_0$) and run for at most H time steps. A steps $H$ steps away from $s_0$ would translate to goals in a goal-oriented MDP. However, it's anytime performance to meet the demands of a dynamic environment is not very good, creating the need for a more sophisticated approach \cite{kolobov2012glutton}.

\subsubsection{LR\textsuperscript{2}TDP}

LR\textsuperscript{2}TDP uses \emph{reverse iterative deepening} and LRTDP-FH to solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$, in that order. It first decides how to act optimally in ($s_0$, 1), equivalent to solving M(1) , then ($s_0$, 2), equivalent to solving $M(2)$, eventually solving $M(H)$ or running out of time halting a $M(h')$ where $h' < H$. This incremental effort allows for reusing state values computed in previous states, so it memorizes all values and solves for smaller horizons by backchaining from the goal LRTDP-FH iterates over search depth while LR\textsuperscript{2}TDP iterates over the distance from the horizon and is likely to be better because:

\begin{itemize}
	\item In many MDPs, the optimal policy $M(h)$ for some $h < H$ is near optimal for M(H), it thus may have a solution when stopped prematurely
	\item it's trials are very short because backchaining from the goal leads to a lots of convergence
	\item It optimally memorizes a large fraction of augmented states to save memory
	\item a large fraction of state space may be unreachable from $s_0$ and it avoid storing information it does not need
	\item it should however use a heuristic to treat states that has never been encountered when stopped prematurely, could be a fall back or a default action.
\end{itemize}

\subsubsection{GLUTTON}

Glutton uses the \emph{Max-Reward} heuristic that relies on knowing the maximum reward $R_{max}$ any action can yield in a state or an upper bound on it. Large branching factors induce that the computation of the Bellman Backups becomes prohibitively expensive. Glutton uses the following mechanisms to optimize for large branching factors:

\begin{itemize}
	\item Subsampling the transition function, is not enough to decrease the amount of computations yet
	\item Separating out Natural Dynamics: to reuse variable samples when generator successors for multiple actions, by separating the actions' effects into natural dynamics (\emph{exogenous}), those due to the action itself (\emph{pure effects}) and those do a mix of both.
	\item Caching the transition function samples: as LR\textsuperscript{2}TDP populates the state-value hash table lazily, Glutton uses  this memory as a cache for samples from the transition function, so it doesn't have to regenerate the action when encountering the same $(s,h)$ pair. When full, it chooses state-action pars for eviction ans replacement randomly.
	\item Default policies: Glutton uses 3 tricks to determine a policy when stopped prematurely. It compares the optimal action from a $h'$ with a simple cyclic policy or a complete random policy and chooses the best of these based on the reward.
\end{itemize}
%---------------------------------------------------------------------
During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 GLUTTON was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, GOURMAND is an online adaptation of the GLUTTON algorithm to test whether LRTDP performs better if it is employed online.

UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

Because real world domains contain Natural Dynamics that introduce large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of state the planner may never reach in a lifetime. Such problems are far better solved using online planners by finding a $\tau$* or an approximation for the current state, executing it and repeating. This is done by choosing a certain lookahead  L and solving MDP \emph{M\textsubscript{s}(L)} to solve \emph{M\textsubscript{0}(H)} where L $<$ H and then select \emph{$\pi$*\textsubscript{L}(s, L)} recommended by the optimal policy. This lookahead value is determined using the per-process time constraint.

\subsubsection{GOURMAND}
In contrast to GLUTTON, GOURMAND always makes an informed choice in any state where it ends up when the time T allocated for solving the MDP runs out because of the online fashion it uses LRTDP. GOURMAND first divides the time among all remaining epochs, determines the largest lookahead L for which it can almost certainly solve any remaining epoch if T is assigned.  It then checks if it can solve even for a larger lookahead L by taking away a small configuration time from future epochs  while still guaranteeing that it can solve for Lookahead L in each of them. This was GOURMAND adaptively picks a good lookahead value at each decision epoch given the per-process constraint T. This can only be achieved by knowing how long it takes for LRTDP to solve a state for certain lookahead values. It uses LR\textsuperscript{2}TDP to collect this data. This version of LRTDP is more efficient because of the early convergences and it allows measuring how long on average the solving of a task takes. GOURMAND uses the same engineering optimizations to to handle evaluate actions, iterating over the successors of s under a by the Bellman Backups. GOURMAND eventually outperforms both GLUTTON and PROST, the winner of the IPPC-11\cite{keller2012prost}.

>>>>

\usepackage{xcolor}
\newcommand\myworries[1]{\textcolor{red}{#1}}
\newcommand\todo[1]{\textcolor{blue}{#1}}

@@

\section{Planners}
\label{sec:planners}

\subsection{FF-replan}
At the first international probabilistic planning competition IPPC-04, most entries were probabilistic planner using Markov Decision Processes. The winner, FF-Replan \cite{FFReplan}, however was based on deterministic planning techniques. FF-Replan is an action selection algorithm for online planning in probabilistic domains which deals with uncertainty by replanning. FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of the planning. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state mapped in an initially empty hash-table. This plan is executed until an unexpected state, not available in the hash-table is encountered.

@@

FF-Replan has some obvious shortcomings.Iit first determinizes the input domain and thus removes all probabilistic information from the problem. The success of the planner is therefore fully reliable on the chose determinization heuristic. As FF-replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-replan performs poorly in domains containing dead-ends. However, by relying on search FF-replan functions with enormous speed and efficiency and caused the International Conference on Automated Planning and Scheduling (ICAPS) to adapt their domains, increasing their \emph{probabilistically interestingness}. Finite horizon MDPs were introduced as an approximation of the infinite optimization problems using a finite horizon. These planning problems accounted for large branching factors as would be caused by \emph{natural dynamics} such as environmental factors and decay. To be successful in these domains planners need to be scalable and sensitive to an expected reward maximization, the evaluation criteria of finite-horizon MDPs.

\subsection{The rise of LRTDP planners}

\myworries{Still need to introduce LRTDPs better and the reason why they emerged}

\emph{Labeled Real-Time Dynamic Programmers} (LRTDPs) operate a series of trials starting from $s_0$.

\todo{Make this list more technical}
\begin{itemize}
	\item Trials start at $s_0$
	\item Choose best action greedily in the current state based on current $V$ value function
	\item Perform Bellman backup on the current state
	\item Sample the outcome as the chosen action
	\item Transition to the new state
	\item Repeat until a goal, dead end or converged state is reached
	\item After trial, special convergence check on all states in the trial is performed to prove the convergence of these values
\end{itemize}

Adapting LRTDP to a finite horizon MDP \emph{M(H)} or \emph{LRTDP-FH} requires for each trial to start at ($s_0$) and run for at most $H$ time steps. The resulting states would translate to the goal states in a goal-oriented MDP. \\

%However, it's anytime performance to meet the demands of a dynamic environment is not very good, creating the need for a more sophisticated approach \cite{kolobov2012glutton}.

A \emph{Labeled Reverse Real-Time Dynamic Programmer} (LR\textsuperscript{2}TDP) uses \emph{reverse iterative deepening} and LRTDP-FH to incrementally solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$ in order by relying on the fact that a successful policy can be produced by solving an MDP-FH $M(H)$ for a smaller horizon $h < H$. A LR\textsuperscript{2}TDP\@ iterates on the distance from the horizon and allows for reusing precomputed state-action plans from converged states. However when dealing with large branching factors Bellman computations tend to become prohibitively expensive. 
\todo{nog een lijstje schrijven van hoe het werkt en waarom er verbetering nodig is}

%\begin{itemize}
%	\item In many MDPs, the optimal policy $M(h)$ for some $h < H$ is near optimal for M(H), it thus may have a solution when stopped prematurely
%	\item it's trials are very short because backchaining from the goal leads to a lots of convergence
%	\item It optimally memorizes a large fraction of augmented states to save memory
%	\item a large fraction of state space may be unreachable from $s_0$ and it avoid storing information it does not need
%	\item it should however use a heuristic to treat states that has never been encountered when stopped prematurely, could be a fall back or a default action.
%\end{itemize}

\subsection{GLUTTON}

\textsc{GLUTTON} introduces an off-line algorithm based on LR\textsuperscript{2}TDP. The \emph{Max-Reward} heuristic used to evaluate \todo{the what?} relies on knowing the maximum reward $R_{max}$ any action can yield in a state or an upper bound on it. In addition to LR\textsuperscript{2}TDP, it introduces the following mechanisms to optimize for large branching factors:

\todo{rewrite this list}
\myworries{is this really necessary?}
\begin{itemize}
	\item Sub sampling the transition function, is not enough to decrease the amount of computations yet
	\item Separating out Natural Dynamics: to reuse variable samples when generator successors for multiple actions, by separating the actions' effects into natural dynamics (\emph{exogenous}), those due to the action itself (\emph{pure effects}) and those do a mix of both.
	\item Caching the transition function samples: as LR\textsuperscript{2}TDP populates the state-value hash table lazily, Glutton uses  this memory as a cache for samples from the transition function, so it doesn't have to regenerate the action when encountering the same $(s,h)$ pair. When full, it chooses state-action pars for eviction ans replacement randomly.
	\item Default policies: Glutton uses 3 tricks to determine a policy when stopped prematurely. It compares the optimal action from a $h'$ with a simple cyclic policy or a complete random policy and chooses the best of these based on the reward.
\end{itemize}

%---------------------------------------------------------------------
%During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 GLUTTON was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, GOURMAND is an online adaptation of the GLUTTON algorithm to test whether LRTDP performs better if it is employed online.
%
%UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

Because real world domains contain Natural Dynamics that introduce large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of states the planner may never reach in a lifetime. Such problems are far better solved using online planners by finding an optimal policy $\tau$* or an approximation of this polict for the \emph{current} state, executing it and repeating. This is done by choosing a certain lookahead value $L$ and solving the MDP \emph{M\textsubscript{s}(L)} to solve \emph{M\textsubscript{0}(H)} where L $<$ H and then select \emph{$\pi$*\textsubscript{L}(s, L)} using the evaluation criteria. This lookahead value is determined using the per-process time constraint.\todo{elaborate on the per-process time constraint}

\subsection{GOURMAND}
\textsc{GOURMAND}, an online adaptation of the \textsc{GLUTTON} planner answers these remarks by adding a lookahead value $L$. \todo{still finish this}

% always makes an informed choice in any state where it ends up when the time T allocated for solving the MDP runs out because of the online fashion it uses LRTDP. GOURMAND first divides the time among all remaining epochs, determines the largest lookahead L for which it can almost certainly solve any remaining epoch if T is assigned.  It then checks if it can solve even for a larger lookahead L by taking away a small configuration time from future epochs  while still guaranteeing that it can solve for Lookahead L in each of them. This was GOURMAND adaptively picks a good lookahead value at each decision epoch given the per-process constraint T. This can only be achieved by knowing how long it takes for LRTDP to solve a state for certain lookahead values. It uses LR\textsuperscript{2}TDP to collect this data. This version of LRTDP is more efficient because of the early convergences and it allows measuring how long on average the solving of a task takes. GOURMAND uses the same engineering optimizations to to handle evaluate actions, iterating over the successors of s under a by the Bellman Backups. GOURMAND eventually outperforms both GLUTTON and PROST, the winner of the IPPC-11\cite{keller2012prost}.

@@
